Constraint Algebras \\ of Multisymplectic Observables

### 2025-01-11 msconstraints-short.tex
Old slides still not working...

%- D0cum3nt ----------------------------------------------------------------------------------------------
\ifHandout
	\documentclass[handout,10pt]{beamer}   
	\setbeameroption{show notes} %print notes   
\else
	\documentclass[10pt]{beamer}
\fi

%- Packages ----------------------------------------------------------------------------------------------
\usepackage{custom-style}
\usepackage{math}
\usepackage{multicol}
\usepackage{xfrac}

@@

\newcommand{\fgmodule}{\mathfrak{X}_{\g}}
\newcommand{\tgtvanform}{I_{\Omega}(N)}
\newcommand{\tgtvf}{\mathfrak X_N(M)}
\newcommand{\vanvf}{I_{\mathfrak{X}}(N)}%{I_N^{\mathfrak X}}
\newcommand{\Ham}{\mathsf{Ham}}
\renewcommand{\ham}{\text{ham}}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{amsmath}
\usepackage{amssymb}

@@

%- T1tle P4g3 -------------------------------------------------------------------------------------------
\title{ Constraint Algebras \\ of Multisymplectic Observables } 
\subtitle{\href{https://sites.google.com/view/xix-yrw-verona/home?authuser=0}{19th International Young Researchers Workshop on Geometry, Dynamics and Field Theory}}
\author[AMM]{
	\href{https://www.antoniomiti.it/}{Antonio Michele Miti}
	{\small joint work w/ Leonid Ryvkin}
}
\institute[SUR]{
	Sapienza Università di Roma \\
	Rome, Italy 	\\
	\vspace{.5em}
  \begin{tabular}[h]{ccc}
      \href{https://www.mat.uniroma1.it/en}{\includegraphics[width=6cm]{./Logos/Sur_logo}} & & 
      \href{https://civis3i.univ-amu.fr/en/civis3i-alliance-programme}{\includegraphics[width=3cm]{Logos/Civis_logo}}
  \end{tabular}    
}
\date[YGMC] % (optional, should be abbreviation of conference name)
{	
	{\vskip 1ex}
	University of Verona, Italy, January 2025
}

%---------------------------------------------------------------------------------------------------------------------------------------------------
%- D0cum3nt ----------------------------------------------------------------------------------------------------------------------------------
\begin{document}
%-------------------------------------------------------------------------------------------------------------------------------------------------
\begin{frame}  % Alternative: \maketitle outside of frame
	\titlepage
	\ifHandout
		\tikz[overlay,remember picture]
		{
	    	%	\node at ($(current page.west)+(1.5,0)$) [rotate=90] {\Huge\textcolor{gray}{\today}};
	    	\node[        
	    		draw,
	    		shape border rotate=90,
			isosceles triangle,
			isosceles triangle apex angle=90,
			fill=yellow]
	        		at ($(current page.north east)-(1,1)$) [rotate=-45] {\textcolor{red}{Handout version}};
		}
	\fi
	\end{frame}
	\addtocounter{framenumber}{-1}
\note{
	%\textbf{\underline{OUTLINE}}:
	%\tableofcontents
	\textbf{Abstract:}
	\\
%
This talk explores key algebraic structures in multisymplectic geometry, a generalization of symplectic geometry where higher-degree $k$-forms replace 2-forms. A central theme is the interplay between symmetries (group actions preserving the differential form) and reduction, which constructs a lower-dimensional space inheriting the geometric structure of the original manifold.

@@

%-------------------------------------------------------------------------------------------------------------------------------------------------
\section{Symplectic Reduction}
\checkpoint	
%-------------------------------------------------------------------------------------------------------------------------------------------------

%-------------------------------------------------------------------------------------------------------------------------------------------------
\begin{frame}{Symplectic geometry (mechanics flavour)}
\begin{columns}[T]
	\begin{column}{.50\linewidth}
		\centering
		\textit{ "geometric approach" to mechanics \dots}
		%
		\begin{columns}
			\begin{column}{.60\linewidth}
				\begin{center}
					\includegraphics[width=0.6\linewidth]{Pictures/pendulum13}			
				\end{center}
			\end{column}	
			\begin{column}{.40\linewidth}
				\begin{center}
					\includegraphics[width=0.45\linewidth]{Pictures/pendulum-phase-space}			
				\end{center}
			\end{column}	
		\end{columns}
		%
		\begin{defblock}[Symplectic Manifold]
			\vspace{-1em}
			\includestandalone[width=1\textwidth]{Pictures/Figure_sym}	
		\end{defblock}
		%
		\pause
		\begin{exblock}[$M = T^\ast Q$ is symplectic]
			with $\omega = d \theta $ given by
			$$ \left.\theta\right\vert_{(q,p)} (v) = p (\pi_\ast v) ~.$$
		\end{exblock}
		%
		\pause
		\vspace{1.2em}
		\centering
		\textit{ based on the notion of \\"states".}		
	\end{column}
	\onslide<1->{\vrule{}}
	\pause
	\begin{column}{.50\linewidth}
		\centering
		\textit{ "algebraic approach" to mechanics \dots}
		\vspace{.5em}	
		\begin{defblock}[Classical Observables]
			Unital, associative, commutative algebra $C^\infty(M)$.
		\end{defblock}
		%
		\vspace{.1em}
		\pause
		\begin{defblock}[Hamiltonian vector fields]
			$\vHam_f \in \mathfrak{X}(M)$ such that:
			$$\iota_{\vHam_f} \omega = -df \quad$$ %$\in B^1(M)$
			%
			\footnotetext{	$\vHam_f$ = \emph{Ham.v.f. pertaining to $f\in C^\infty(M)$}.}
		\end{defblock}
		\vspace{.1em}
		%
		\begin{defblock}[Poisson Algebra of Observables]
			$C^\infty(M)$ is a Poisson algebra with
			$$\{f,g\} = \iota_{\vHam_g} \iota_{\vHam_f} \omega = \omega(\vHam_f,\vHam_g) ~.$$
		\end{defblock}
		%
		\pause
		\vspace{.15em}
		\centering
		\textit{ based on the notion of \\"measurable quantities".}						
	\end{column}
\end{columns}
\end{frame}
\note[itemize]{
	\footnotesize

	\item We work in the framework of multisymplectic geometry which is one of the possible generalizations of the well-established field of symplectic geometry.

	\item To recall what symplectic geometry is let me assume a particular point of view: mechanics.
	\\
	Idea:"
	Symplectic geometry is a branch of differential geometry studying symplectic manifolds; it originated as a formalization of the mathematical apparatus of classical mechanics and geometric optics."{\href{https://ncatlab.org/nlab/show/symplectic+geometry}{nlab}}

	Namely, a sym. mfd. is the geometric structure encoding the phase space of conservative, ordinary, classical, mechanical systems.

	\item $\theta$ = \emph{tautological 1-form}.
		$\theta$ evaluated at $p\in T^*Q$ in the fibre of $q\in Q$ and contracted with $v$ coincides with the form $p$ evaluated at $q$ and contracted with the push forward of $v$.

	\item We identify a special class of vector fields.
		Out of them one can define a Lie bracket.

	\item Poisson is a Lie algebra with the extra property of compatibility with the associative product (Leibniz rule)

	\item take away message: geometric (based on "states") vs algebraic (based on "measurable quantities").7
}
%-------------------------------------------------------------------------------------------------------------------------------------------------

%------------------------------------------------------------------------------------------------
\begin{frame}{Reminder: momentum maps in symplectic geometry}\label{frame:symplecticmomaps}
	Consider $\theta:G\curvearrowright M$ symplectic,~~ $\underline{\cdot}:\mathfrak{g}\to \mathfrak{X}(M)$ infinitesimal action.
	\vfill

	\begin{columns}[T]
		\setlength{\belowdisplayskip}{5pt}
		\begin{column}{.50\linewidth}
			%
			\centering \it
			\onslide<2->{
				\begin{defblock}[Equivariant moment map]
					Smooth map $$\mu:M\to \mathfrak{g}^\ast$$
					such that:
					\begin{itemize}
						\item[i.] $d\langle \mu,\xi\rangle = -\iota_{\underline{\xi}}\omega$ 
						~\qquad, $\forall \xi \in \mathfrak{g}$
						\item[ii.] $\mu \circ \theta_g = Ad_g^\ast \circ \mu$
						 \qquad\small, $\forall g \in G$
					\end{itemize}
				\end{defblock}
			}
		\end{column}	
		%
		\onslide<2->{\vrule{}}
		%
		\begin{column}{.50\linewidth}
			\onslide<3->{			
				\begin{defblock}[Comoment map]
					Linear map $$\widetilde{\mu}: \mathfrak{g}\to C^\infty(M,\omega)$$
					such that:
					\begin{itemize}
						\item[i.] $d\widetilde{\mu}(\xi) = -\iota_{\underline{\xi}}\omega$ 
						\qquad~\small, $\forall \xi \in \mathfrak{g}$
						\item[ii.] $\widetilde{\mu}([\xi,\eta]) = \lbrace\widetilde{\mu}(\xi),\widetilde{\mu}(\eta)\rbrace$ \small, $\forall \xi,\eta \in \mathfrak{g}$
					\end{itemize}
				\end{defblock}
			}
		\end{column}	
	\end{columns}	
	%
	\pause
	\vspace{1em}
	%
	\begin{columns}[]
		\setlength{\belowdisplayskip}{5pt}
		\begin{column}{.40\linewidth}
			%
			\centering \it
			\onslide<5->{
				\begin{upshotblocktitle}[Duality]
					\begin{displaymath}
						\mu(x) : \mathfrak{\xi} \mapsto \widetilde{\mu}(\xi)\big\vert_x
					\end{displaymath}
					%
					\emph{
					\small
					"duality wrt. the currying operation"					
					}
				\end{upshotblocktitle}
			}
		\end{column}	
		%
		%
		\begin{column}{.60\linewidth}
			\onslide<4->{			
				\begin{upshotblocktitle}[$\widetilde{\mu}$ as a lift]
					\begin{displaymath}
						\begin{tikzcd}[ampersand replacement = \&]
						 \& C^\infty(M,\omega) \ar[d,"\vHam"]
						 \\
						 \mathfrak{g} \ar[ur,dashed,sloped,"\widetilde{\mu}"]\ar[r,"\underline{\cdot}"'] \& \mathfrak{X}(M)
						\end{tikzcd}
					\end{displaymath}
					%
					\emph{
					\small
					"it is a lift (in the Lie category) of the infinitesimal action by the assigment of hamiltonian v.fields."					
					}
				\end{upshotblocktitle}
			}
		\end{column}	
	\end{columns}		
\end{frame}
\note[itemize]{
	\item consider an action preserving the symplectic form.
	\item[-] $\langle \cdot, \cdot \rangle$ is the natural pairing of $\mathfrak{g}$ and $\mathfrak{g}^\ast$.
		\item[-] $\theta_g$ is the diffeomorphism associated to $g\in G$ via $\theta:G\curvearrowright M$
		\item[-] $Ad^\ast_g$ is the coadjoint action $G\curvearrowright \mathfrak{g}^\ast$
	\item The moment map can be reexpressed as a comomentum map.
	\item Observe that ii. on the left always implies ii. on the right. The converse is trure only if $G$ is connected.
	\item $G\curvearrowright M$ is said "Hamiltonian" (see slide \ref{frame:gisthamaction} in appendix) iff $\exists$ comoment map $\widetilde{\mu}$.
	\item \emph{moment map: Describe ${\color{orange}G\curvearrowright M}$ in terms of $\color{blue}C^\infty(M)\to\X(M)$.}
	\begin{center}
	\begin{tikzpicture}[scale=1.5]
		\node[orange] (A) at (0,0) {$\g$};
		\node[gray] (A1) at (0,-.3) {$\xi$};
		\node[blue] (B) at (1,1) {$C^\infty(M)$};
		\node[gray] (B2) at (1.55,1) {$f$};
		\node[orange] (C) at (1,0) {$\X(M)$};
		\node[gray] (C1) at (1,-.3) {$\underline\xi$};
		\node[gray] (C2) at (1.55,0) {$X_f$};

		\path[->,blue,dashed] (A) edge node[above left] {$\langle\mu,\cdot\,\rangle$} (B);
		\path[->,orange] (A) edge (C);
		\path[->] (B) edge (C);
		\path[|->,gray] (A1) edge (C1);
		\path[|->,gray] (B2) edge (C2);

		\begin{scope}[xshift=3cm]
			\node at (-.32,.8) {as Lie algebras};
			\node at (-.13,.2) {i.e., {\color{blue}$\mu_\xi$} generates {\color{orange}$\underline\xi$}};
		\end{scope}
	\end{tikzpicture}

		\footnotesize		(\emph{moment map:} ${\color{blue}\mu:M\to\g^*}$, 
	 \emph{Hamiltonian $G$-space:} 	$(M,\omega,G,\mu)$)
	\end{center}
	\item
	 $\alpha\mapsto X_\alpha$ is a homomorphism of Leibniz algebras:
	$$\d\{\alpha,\beta\} = \d\L_{X_\alpha}\beta = \L_{X_\alpha}\iota_{X_\beta}\omega = \iota_{[X_\alpha,X_\beta]}\omega ~{\color{black!80} \implies}~
	X_{\{\alpha,\beta\}} = [X_\alpha,X_\beta]$$
}
%------------------------------------------------------------------------------------------------

\subsection{Regular reduction}
%------------------------------------------------------------------------------------------------
\begin{frame}{Regular reduction in symplectic geometry}
	\textbf{\color{UniGreen}Symplectic reduction:}~~
	\\ 
	{\it \small
	Procedure associating to any (suitably regular) pair of symplectic manifold and Hamiltonian action another symplectic manifold of smaller dimension.}
	\vfill
	\pause
	\begin{thmblock}[Marsden-Weinstein reduction \cite{MarsdenWeinstein74}]
		\vspace{-.4em}\hspace{-1em}
		\begin{tabular}{l p{14cm}}
		    Given: & $(M,\omega)$ symplectic
		    \\
		    & $G\curvearrowright M$ symplectic with equivariant momap. $\mu:M\to \mathfrak{g}^*$
		    \\[.2em]
		    Assume: & $\phi \in \mathfrak{g}^*$ regular value of $\mu$ 
		    \qquad\quad \footnotesize \textcolor{gray}{($\Rightarrow$ $\mu^{-1}(\phi)\hookrightarrow M$ smooth embedding)}
		    \\
			& $G_\phi\curvearrowright \mu^{-1}(\phi)$ free and proper
			\quad \footnotesize \textcolor{gray}{($\Rightarrow$ $\mu^{-1}(\phi)/G_\phi$ smooth manifold)}
			\\[.4em]
			Then: & $\exists!$ symplectic structure $\omega_\phi$ in $M_\phi:= \mu^{-1}(\phi)/G_\phi$ \\
			& s.t. $\pi^\ast \omega_\phi = j^\ast \omega$ 
			\qquad {\footnotesize with $j:\mu^{-1}(\phi) \hookrightarrow M$ and $\pi:\mu^{-1}(\phi)\twoheadrightarrow M_\mu$}
		\end{tabular}
		\vspace{-.4em}
	\end{thmblock}
	%
	\vfill
	\pause
	\begin{columns}
		\begin{column}{0.40\textwidth}
			\includegraphics[width=\textwidth]{Pictures/Reduction}
		\end{column}

		\begin{column}{0.6\textwidth}
				\textbf{\color{UniGreen}In mechanics:}~~
				\\
			{\it \small
				it embodies the process of restricting the dynamics of the system to the level sets of the conserved quantities pertaining to the symmetry group.		
			}
			\\[.2em]
			\color{gray}\small( e.g. restricting to studying a point-like particle in a central potential by studying it in radial coordinates)
		\end{column}	
	\end{columns}	
\end{frame}
\note[itemize]{
	\item {Symplectic Reduction} takes two inputs:
	\\ 1. Hamiltonian $G$-space $(M,\omega,{\color{orange}G},{\color{blue}\mu})$
	\\ 2. parameter ${\color{blue}\phi}\in\g^*$
	\item The \emph{reduced space} is $M_\phi:={\color{blue}\mu^{-1}(\phi)}{\color{orange}/G_\phi}$.
	\item (Marsden--Weinstein '74, Meyer '73) THM:\\
		If $\mu^{-1}(\phi)\subset M$ is smooth and $G_\phi\curvearrowright \mu^{-1}(\phi)$ is free and proper, then there is a \textbf{unique symplectic form} $\omega_\phi\in\Omega^2(M_\phi)$ such that $\pi^*\omega_\phi = i^*\omega$.

	\begin{center}
	\begin{tikzpicture}[scale=2]
		\node[blue] (A) at (0,0) {$\mu^{-1}(\phi)$};
		\node[gray] (A1) at (0,.3) {$i^*\omega$};
		\node[gray] (A2) at (-.6,0) {$\pi^*\omega_\phi$};
		\node[blue] (B) at (1,0) {$M$};
		\node[gray] (B1) at (1,.3) {$\omega$};
		\node[orange] (C) at (0,-.8) {$M_\phi$};
		\node[gray] (C2) at (-.6,-.8) {$\omega_\phi$};

		\path[right hook->,blue] (A) edge node[above] {$i$} (B);
		\path[orange,->] (A) edge node[left] {$\pi$} (C);
		\begin{scope}[xshift=-1.8cm]
			\node[blue] at (0,0) {\small restrict to $\{\mu=\phi\}$};
			\node[orange] at (.15,-.4) {\small quotient by $G_\phi$};
		\end{scope}
	\end{tikzpicture}
	\end{center}
	\item Heuristic Approach to Reduction:
	\\1. describe $G\curvearrowright M$ in terms of $\omega$ 	\hspace{1cm}
	({\color{green}moment map $\mu$})
	\\2. identify a distinguished reduced space \hspace{1cm}
	({\color{green}reduction at $\mu=0$})
	\\3. use the ambiguity in 1.\ to obtain a family of reduced spaces \hspace{1cm}
	({\color{green}reduction at $\mu-\phi=0$, i.e.\ reduction at $\mu=\phi$}
	\\
	{\tiny Note: If $G_\phi\neq G$, then $\mu-\phi:M\to\g^*$ is not a moment map for either $G\curvearrowright M$ or $G_\phi\curvearrowright M$.}

}
%------------------------------------------------------------------------------------------------

%-------------------------------------------------------------------------------------------------------------------------------------------------
\subsection{Singular reduction}
\begin{frame}{Symplectic singular reduction schemes}
	\begin{block}{The gist of singular reduction}
 		\begin{itemize}
 			\item[-] when $\mu$ is singular (i.e. $\mu^{-1}(0)$ is not a mfd.), the (geometrically) reduced space may not exist.
 			\item[-] a \emph{singular reduction scheme} is a procedure to construct a "reduced" algebra of observable out of the given data
 			\item[-] such that it corresponds to the algebra of observable of the reduced manifold in the regular case.
 		\end{itemize}
	\end{block}
 	%
 	\pause
 	%
	\begin{thmblock}[Sniatycki-Weinstein reduction \cite{SniatyckiWeinstein83}]
		\vspace{-.4em}\hspace{-1em}
		\begin{tabular}{l p{14cm}}
		    Given: & $(M,\omega)$ symplectic
		    \\
		    & $G\curvearrowright M$ symplectic with equivariant momap. $\mu:M\to \mathfrak{g}^*$
			\\[.4em]
			Then: & 
			$\displaystyle \left[\sfrac{C^\infty(M)}{I_\mu}\right]^G$
			admits a Poisson algebra structure 
			\blfootnote{$I_\mu$ = associative ideal generated by $\widetilde{\mu}(\g)$}			
			\\
			& it agrees with the M--W reduction in the regular case.		
		\end{tabular}
		\vspace{-.4em}
	\end{thmblock}

\end{frame}
\note[itemize]{
 \item Let $(M,\omega,G,\mu)$ be a symplectic Hamiltonian $G$-space.
 \item The \emph{momentum ideal} is the associative ideal $I_\mu\subset C^\infty(M)$ generated by the momenta $\mu_{\xi}$ for any $\xi\in \mathfrak{g}$.
		Namely
		\[
			I_\mu 
			= 
			\Big\langle \mu_\xi \Big\rangle_{\xi \in \mathfrak{g}}^{\text{\tiny asso.}} 
			=
			\left\{
				\left.
					\sum_{i=1}^n f_i ~ \mu_{\xi_i}
				\right|~
					f_i \in C^\infty(M),~ \xi_i\in \mathfrak{g},~  1 \leq i \leq n
			\right\}
		\]

 \item The \emph{\'Sniatycki--Weinstein reduction} is the Poisson algebra
		\[
			\left(C^\infty(M)/I_\mu\right)^G.
		\]

}
%-------------------------------------------------------------------------------------------------------------------------------------------------

%-------------------------------------------------------------------------------------------------------------------------------------------------
\section{Multisymplectic Geometry}
\checkpoint	
%-------------------------------------------------------------------------------------------------------------------------------------------------

%-------------------------------------------------------------------------------------------------------------------------------------------------
\begin{frame}{From symplectic to {multi}symplectic} 
	%
	\begin{center}
		$-$ \emph{multisymplectic means \textbf{going higher} in the degree of $\omega$} $-$
	\end{center}
	\pause
	\begin{defblock}[$n$-plectic manifold ~\emph{(Cantrijn, Ibort, De Le\'on)} \cite{Cantrun2017}]
		\includestandalone[width=0.95\textwidth]{Pictures/Figure_multisym}	
	\end{defblock}
	%
	\pause
		\begin{table}
			\begin{tabular}{c c c}
				symplectic forms \small($n=1$) & $\leftrightsquigarrow$ & volume forms \small($n= \text{dim}(M)-1$)
			\end{tabular}
		\end{table}

	\vfill
	\pause
	\begin{block}{Historical motivation}
		Mechanics: geometrical foundations of \textit{(first-order)} field theories.
	\end{block}
	%
	\begin{table}
		\ifHandout
		%
		\else
		\only<4>{
		\begin{tabular}{|p{0.2\textwidth}|p{0.3\textwidth}|p{0.35\textwidth}|} 
            \hline
            \parbox[][20pt][c]{0.2\textwidth}{mechanics} & \multicolumn{2}{c|}{geometry} \\
            \hline
            \parbox[][20pt][c]{0.2\textwidth}{phase space} & symplectic manifold &  \\[.25em]
            \parbox[][20pt][c]{0.2\textwidth}{classical \\ observables} & Poisson algebra &  \\[.25em]
            \parbox[][20pt][c]{0.2\textwidth}{symmetries} &  group actions admitting comoment map &  
            \\
            \hline
  \multicolumn{1}{c}{}
            &  \multicolumn{1}{@{}c@{}}{$\underbrace{\hspace*{.3\textwidth}}_{\text{point-like particles systems}}$} 
            &  \multicolumn{1}{@{}c@{}}{}              \\
		\end{tabular}
		}
		\fi
		\onslide<5->{
		\begin{tabular}{|p{0.2\textwidth}|p{0.3\textwidth}|p{0.35\textwidth}|} 
            \hline
            \parbox[][20pt][c]{0.2\textwidth}{mechanics} & \multicolumn{2}{c|}{geometry} \\
            \hline
            \parbox[][20pt][c]{0.2\textwidth}{phase space} & symplectic manifold & multisymplectic manifold \\[.25em]
            \parbox[][20pt][c]{0.2\textwidth}{classical \\ observables} & Poisson algebra & $L_\infty$-algebra \\[.25em]
            \parbox[][20pt][c]{0.2\textwidth}{symmetries} &  group actions admitting comoment map & group actions admitting (homotopy) comomentum map
            \\
            \hline
  			\multicolumn{1}{c}{}
            &  \multicolumn{1}{@{}c@{}}{$\underbrace{\hspace*{.3\textwidth}}_{\text{point-like particles systems}}$} 
            &
            \multicolumn{1}{@{}c@{}}{$\underbrace{\hspace*{.3\textwidth}}_{\text{field-theoretic systems}}$} 
               \\
		\end{tabular}
		}
	\end{table}

\end{frame}
\note[itemize]{
	\item Historically, the interest in multisymplectic manifolds, has been motivated by the need for understanding the geometrical foundations of first-order classical field theories.
	The key point is that, just as one can associate a symplectic manifold to an ordinary classical mechanical system (e.g. a single
point-like particle constrained to some manifold), it is possible to associate a multisymplectic
manifold to any classical field system (e.g. a continuous medium like a filament or a fluid). See frame Extra-\ref{Frame:Ms-Field-Mechanics}

	\item General ideas basic parallelisms with caveats
	\item caveat: points in multiphase spaces are not states
	\item the table hides the duality between geometric and algebraic approaches to the problem.
	\item	Mechanics: geometrical foundations of \textit{(first-order)} field theories.
		\begin{itemize}
		 \item[-] Kijowski, W. Tulczyjew \cite{Kijowski1979}; %(1979)
		 \item[-] Cariñena, Crampin, Ibort \cite{Carinena1991b};% (1991)
		 \item[-] Gotay, Isenberg, Marsden, Montgomery \cite{Gimmsy1};%(1998)
		 \\ $\cdots$
		\end{itemize}
}
%-------------------------------------------------------------------------------------------------------------------------------------------------

%-------------------------------------------------------------------------------------------------------------------------------------------------
\begin{frame}{Observables in $n$-plectic geometry}
	%
	\begin{defblock}[Hamiltonian $(n-1)$-forms]
		\begin{displaymath}
			\Omega^{n-1}_{ham}(M,\omega) 	:=
			\biggr\{ \sigma \in  \Omega^{n-1}(M) \; \biggr\vert \; 
				\exists \vHam_\sigma \in \mathfrak{X}(M) ~:~ 
				\tikz[baseline,remember picture]{\node[rounded corners,
                        fill=orange!5,draw=orange!30,anchor=base]            
            			(target) {$d \sigma = -\iota_{\vHam_\sigma} \omega$ };
            	}				
				~\biggr\} 
			\end{displaymath}
	\end{defblock}

	\pause
		\tikz[overlay,remember picture]
		{
			\node[rounded corners,
                 fill=orange!5,draw=orange!30,anchor=base]
            	 (base) at ($(current page.north east)-(2,1)$) [rotate=-0,text width=3.5cm,align=center] {\footnotesize{\textcolor{red}{Hamilton-DeDonder-Weyl \\equation}}};
		}	
	\begin{tikzpicture}[overlay,remember picture]
    	\path[->] (base.south) edge[bend right,red](target.north);
    \end{tikzpicture}
	%
	\vfill
	\begin{columns}[T]
		\pause
		\setlength{\belowdisplayskip}{5pt}
		\begin{column}{.50\linewidth}
			%
			\vspace{-1em}
			\begin{thmblock}[Observables $L_\infty$-algebra]
				$\Omega^{n-1}_{ham}(M,\omega)$ endowed with
				\vspace{-.5em}
				\begin{displaymath}
					\lbrace \sigma_1, \sigma_2 \rbrace =			
					~ - \iota_{\vHam_1}\iota_{\vHam_2} \omega 
				\end{displaymath}			
				can be "completed" to a \\ $L_\infty-algebra$.
			\end{thmblock}
			%	
		\end{column}	
		%
		\pause
		\begin{column}{.50\linewidth}
		%
			\begin{itemize}
				\item[\cmark] Skew-symmetric;
				\item[\xmark] multiplication of observables;
				\item[\xmark] Jacobi equation;
				%\\ \hspace*{4.25em} full-fledged Jacobi equation;
				\item[\smark] Jacobi equation \emph{up to homotopies}.
			\end{itemize}			
			%
				\[
					\small
					\{\alpha,\{\beta,\gamma\}\} + \text{\it cyc. perm.}
					= {\color{orange}\d \iota_{X_\alpha} \iota_{X_\beta} \iota_{X_\gamma}\omega}
				\]
		\end{column}	
	\end{columns}
	%
	\vfill
	\pause
	Interesting alternatives:
	\begin{itemize}
		\item  descend to $\Omega^{k-1}_{ham}(M) / {\color{orange}\d\hspace{1pt}\Omega^{k-2}(M)}$, hence getting a Lie algebra;
		\item Incorporate the {\color{orange}discrepancy} as part of the data of the space of observables $ 
				L_\infty(M,\omega) = \Omega_{ham}^{k-1}(M)\oplus{\color{orange}\Omega^{\leq k-2}(M)}$
			%a \textbf{homotopy Lie algebra} or \textbf{$L_\infty$-algebra}.
	\end{itemize}

\end{frame}
\note[itemize]{
	\item
}
%-------------------------------------------------------------------------------------------------------------------------------------------------

%---------------------------------------------------------------------------------------------------------------------------------------------------
\subsection{$L_\infty$-algebra of Observables}
\begin{frame}[fragile,t]{$L_\infty$-algebra of Observables (higher observables) }
	Let be $(M,\omega)$ a $n$-plectic manifold.
	\begin{defblock}[$L_\infty$-algebra of observables ~\emph{(Rogers)} ~\cite{Rogers2010}]

		\hspace{.25em} $L\infty(M,\omega)$ is given by:

		\begin{itemize}
			\item[•] a cochain-complex $(L,\{\cdot\}_1)$ 
		\end{itemize}
		\begin{center}
		\ifHandout
			\includestandalone{Pictures/Figure_Observables}	
		\else
			\includestandalone{Pictures/Frame_Observables}
		\fi				
		\end{center}
		\onslide<2->{
			\begin{itemize}
				\item[•] with $n$ (skew-symmetric) multibrackets $(2 \leq k \leq n+1)$
			\end{itemize}
			\begin{center}
				\includestandalone{Pictures/Equation_Multibracket}	
			\end{center}
		}
		%
	\end{defblock}
  \end{frame}
 \note[itemize]{
	\item if symplectic manifolds are the symmetric take on mechanics, Poisson algebras are the algebraic counterpart.
 	\item A Lie algebra is associated to an ordinary symplectic manifold (its Poisson algebra).
	%(Underlying this is a Lie algebra, whose Lie bracket is the Poisson bracket.)
	Similarly, one associates an Lie-$n$ algebra to any $n$-plectic manifold.
 	% https://ncatlab.org/nlab/show/n-plectic+geometry 	 
 	 %https://ncatlab.org/nlab/show/Poisson+bracket+Lie+n-algebra
	 \item Basically, the higher observables algebra is a chunk of the de Rham complex of $M$ with inverted grading( convention employed here) and an extra structure called "multibrackets".
 	\item ( In the 1-plectic case it reduces to the corresponding Poisson algebra of classical observables)
 	\item Rogers associated to any n-plectic mfd a $L\-\infty$ algebra, Zambon generalized it to the pre-n-plectic case.
 	\item Recognize in the definition of $\{\cdot,\ldots,\cdot\}_k$ the contraction with hamiltonian fields $v_\sigma$ w.r.t. $\sigma$.
  	\item Note $	\iota_{v_{\sigma_1}}\cdots\iota_{v_{\sigma_k}} = (-)^{(k-1)+(k-2)+\dots+1}\iota_{v_{\sigma_k}}\cdots\iota_{v_{\sigma_1}} = (-)^{\frac{k(k-1)}{2}}\iota_{v_{\sigma_k}}\cdots\iota_{v_{\sigma_1}}$ 
 	The definition usually find in literature of Rogers multibrackets involves the coefficient $ (-)^{\frac{k(k-1)}{2}} = -\varsigma(k-1) = (-)^{k+1} \varsigma(k)$.
  \item higher observables is Special instance of a more general object  called $L\-\infty$ Algebra...
 }
%------------------------------------------------------------------------------------------------

%-------------------------------------------------------------------------------------------------------------------------------------------------
\section{Multisymplectic observables reduction}
\checkpoint	
%-------------------------------------------------------------------------------------------------------------------------------------------------
%-------------------------------------------------------------------------------------------------------------------------------------------------
\begin{frame}{Singular Reduction: Roadmap}
	%
	\begin{block}{Data:}
			\begin{itemize}
				\item A \emph{constraint set} $N$ (possibly singular),
				\item An infinitesimal action preserving $N$.
			\end{itemize}
	\end{block}
	%
	\vfill
	\pause
	%
	\begin{block}{Goal:}
		\begin{itemize}
			\item Obtain a "reduced" observables algebra out of the data.
		\end{itemize}
	\end{block}
	%
	\vfill
	\pause
	%
	\begin{block}{Strategy:}
		\begin{enumerate}
			\item Define smooth fields/forms \emph{tangent to $N$},
			\item define smooth fields/forms \emph{vanishing along $N$},
			\item define \emph{reducible fields} requiring the preservation of the vanishing objects,
			\item define \emph{reducible forms} requiring their conservation w.r.t. the infinitesimal action,
			\item define \emph{reducible and vanishing observables},
			\item \textbf{quotient}
		\end{enumerate}
	\end{block}

\end{frame}
\note[itemize]{
	\item
}

%-------------------------------------------------------------------------------------------------------------------------------------------------

%-------------------------------------------------------------------------------------------------------------------------------------------------
\begin{frame}[shrink]{Smooth objects on a singular set}
	Consider $N$ closed subset of $M$.
	\vfill
	\pause
	\begin{defblock}
	 $I_N$ = ideal of smooth functions vanishing over $N$.
	\end{defblock}
	\vfill
	\pause

>>>>

%- Document ----------------------------------------------------------------------------------------------
\ifHandout
	\documentclass[handout,10pt]{beamer}   
	\setbeameroption{show notes} %print notes   
\else
	\documentclass[10pt]{beamer}
\fi

%- Packages ----------------------------------------------------------------------------------------------
\usepackage{custom-style}
\usepackage{math}
\usepackage{multicol}
\usepackage{xfrac}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{tikz}
\usetikzlibrary{backgrounds}

@@

\newcommand{\fgmodule}{\mathfrak{X}_{\g}}
\newcommand{\tgtvanform}{I_{\Omega}(N)}
\newcommand{\tgtvf}{\mathfrak X_N(M)}
\newcommand{\vanvf}{I_{\mathfrak{X}}(N)}%{I_N^{\mathfrak X}}
\newcommand{\Ham}{\mathsf{Ham}}
\renewcommand{\ham}{\text{ham}}

@@

%- Title Page -------------------------------------------------------------------------------------------
\title{ Constraint Algebras \\ of Multisymplectic Observables } 
\subtitle{\href{https://sites.google.com/view/xix-yrw-verona/home?authuser=0}{19th International Young Researchers Workshop on Geometry, Dynamics and Field Theory}}
\author[AMM]{
	\href{https://www.antoniomiti.it/}{Antonio Michele Miti}
	{\small joint work w/ Leonid Ryvkin}
}
\institute[SUR]{
	Sapienza Università di Roma \\
	Rome, Italy 	\\
	\vspace{.5em}
  \begin{tabular}[h]{ccc}
      \href{https://www.mat.uniroma1.it/en}{\includegraphics[width=6cm]{./Logos/Sur_logo}} & & 
      \href{https://civis3i.univ-amu.fr/en/civis3i-alliance-programme}{\includegraphics[width=3cm]{Logos/Civis_logo}}
  \end{tabular}    
}
\date[YGMC] % (optional, should be abbreviation of conference name)
{	
	{\vskip 1ex}
	University of Verona, Italy, January 2025
}

%---------------------------------------------------------------------------------------------------------------------------------------------------
%- Document ----------------------------------------------------------------------------------------------------------------------------------
\begin{document}
%-------------------------------------------------------------------------------------------------------------------------------------------------
\begin{frame}  % Alternative: \maketitle outside of frame
	\titlepage
	\ifHandout
		\tikz[overlay,remember picture]
		{
	    	%	\node at ($(current page.west)+(1.5,0)$) [rotate=90] {\Huge\textcolor{gray}{\today}};
	    	\node[        
	    		draw,
	    		shape border rotate=90,
			isosceles triangle,
			isosceles triangle apex angle=90,
			fill=yellow]
	        		at ($(current page.north east)-(1,1)$) [rotate=-45] {\textcolor{red}{Handout version}};
		}
	\fi
	\end{frame}
	\addtocounter{framenumber}{-1}
\note{
	%\textbf{\underline{OUTLINE}}:
	%\tableofcontents
	\textbf{Abstract:}
	\\
%
This talk explores key algebraic structures in multisymplectic geometry, a generalization of symplectic geometry where higher-degree $k$-forms replace 2-forms. A central theme is the interplay between symmetries (group actions preserving the differential form) and reduction, which constructs a lower-dimensional space inheriting the geometric structure of the original manifold.

@@
